---
title: "Lab 03 - Nobel laureates"
author: "Lily Botha"
date: "02/01/2026"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

Here is a link to the [lab instructions](https://datascience4psych.github.io/DataScience4Psych/lab03.html).

### Exercise 1

There are 935 observations and 26 variables. Each row represents a person/laureate. 

### Exercise 2

```{r nobel_living}
nobel_living <- nobel %>% 
  filter(is.na(died_date), 
         !is.na(country), 
         gender != "org"
  )
```

### Exercise 3

It looks like most Nobel laureates were based in the US when they won their prize, but physics, medicine, and chemistry awards are almost evenly split between the US and other countries. The plot looks very similar to the Buzzfeed article.

```{r country_us}
#create new variable based on if statement: if country = "USA", set country_us to "USA", if not, set to "Other"
nobel_living <- nobel_living %>% 
  mutate(country_us = if_else(country == "USA", "USA", "Other")
  ) 
```

```{r science}
#filter by observations within a specific column (category)
nobel_living_science <- nobel_living %>% 
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics")
  ) 
```

```{r bar_plot}
nobel_living_science %>% 
  ggplot(aes(x = category, fill = country_us)) + 
  geom_bar(position = "fill") + 
  labs(title = "Category of Nobel Prize vs. Location Won", 
       subtitle = "where Nobel laureates were based when they won", 
       x = "Category", 
       y = "Frequency", 
       fill = "Country"
  ) + 
  coord_flip()
```

### Exercise 4

105 living Nobel prize winners were born in the US. 

```{r born_country_us}
born_country_us <- nobel_living_science %>% 
  mutate(born_country_us = if_else(born_country == "USA", "USA", "Other")
  )
```

```{r count}
born_country_us %>% count(born_country_us)
```

### Exercise 5

```{r add_to_bar_plot}
# still need to fix this
born_country_us %>% ggplot() + geom_bar(aes(x = country_us, fill = born_country_us), position = "fill") + facet_wrap(~category)
```

...

### Exercise 6

...
